\ \\ \LARGE\bf Speculative Evaluation in Particle Swarm Optimization \thanks{Matthew Gardner, Andrew McNabb, and Kevin Seppi are with the Department of Computer Science, Brigham Young University, 3361 TMCB, Provo, UT 84602 (phone: 801-422-8717; email: \{mjg82,a,k\}@cs.byu.edu).}

### 2010-03-15 honors_thesis/journal_paper/paper.tex

Previous parallel techniques use additional processors to increase the
number of particles in the swarm, either by adding individual particles or by
adding entire new sub-swarms but in all cases increasing the number of particles.  The number of iterations of PSO that the
algorithms can perform is thus inherently limited by the time it takes to
evaluate the objective function---additional processors add more particles, not
more iterations.
for many functions there comes a point of diminishing returns with respect to adding particles.
>>>>>>>>> TODO WARNING WARNING MADE UP NUMBERS GET REAL ONES!! <<<<<<<<<<<<<<
For the well-know benchmark function Sphere and as shown in Figure XXXXa increasing the swarm size from 20 to 120 has a significant effect on the 
speed of convergence. In this case, the swam goes below $1*10^5$ in x iteration with 20 particles, but 
achieves this same value in  y iterations with a swarms size of 120, the additional 100 particles.
However consider the change going from 2000 particles to 2100 particles. Here the additional 100 particles
reduce the number of iterations to achieve $1*10^5$ from x iterations to y.
Other benchmark function behave similarly. ans shown in fig xxxxb for ?? and xxxxc ?? (in all case 20 dimensions, average of 50 runs).
Certainly these benchmark functions do not prove that this will be the case for all functions,
but it does provide some evidence that for at least some functions, adding additional particles does not seem to have much benefit
after some point, some time earlier fig xx some time later. This phenomena has been commented on by others
in the form of recommended swarms sizes for various types of functions cite standard swarm paper.
\begin{quote}
Skip this??

@@

There are many ways of defining the neighbors of a particle.  Different
neighborhood topologies have varying performance on different kinds of
functions, as would be expected by the No Free Lunch
theorem~\cite{wolpert-tec97}.  An in-depth discussion of topologies in PSO is
beyond the scope of this paper.  The two most common topologies used in the
literature are the Ring topology and the Complete
topology~\cite{bratton-sis07}.  In this paper we use those topologies, as well
as a parallel surrogate for the Complete topology called Random that has been
shown to simulate the behavior of Complete with far less
communication~\cite{mcnabb-cec09}.  Both the Ring and the Random topologies can
be defined as having $n$ neighbors; we use 2 neighbors for all of our
experiments.
%TODO except as noted??

>>>>

Previous parallel techniques use additional processors to increase the
number of particles in the swarm, either by adding individual particles or by
adding entire new sub-swarms but in all cases increasing the number of particles.  The number of iterations of PSO that the
algorithms can perform is thus inherently limited by the time it takes to
evaluate the objective function---additional processors add more particles, not
more iterations.
for many functions there comes a point of diminishing returns with respect to adding particles.
!!!! TODO WARNING WARNING MADE UP NUMBERS GET REAL ONES!! !!!!!
For the well-know benchmark function Sphere and as shown in Figure XXXXa increasing the swarm size from 20 to 120 has a significant effect on the 
speed of convergence. In this case, the swam goes below $1*10^5$ in x iteration with 20 particles, but 
achieves this same value in  y iterations with a swarms size of 120, the additional 100 particles.
However consider the change going from 2000 particles to 2100 particles. Here the additional 100 particles
reduce the number of iterations to achieve $1*10^5$ from x iterations to y.
Other benchmark function behave similarly. ans shown in fig xxxxb for ?? and xxxxc ?? (in all case 20 dimensions, average of 50 runs).
Certainly these benchmark functions do not prove that this will be the case for all functions,
but it does provide some evidence that for at least some functions, adding additional particles does not seem to have much benefit
after some point, some time earlier fig xx some time later. This phenomena has been commented on by others
in the form of recommended swarms sizes for various types of functions cite standard swarm paper.
\begin{quote}
Skip this??

@@

There are many ways of defining the neighbors of a particle.  Different
neighborhood topologies have varying performance on different kinds of
functions, as would be expected by the No Free Lunch
theorem~\cite{wolpert-tec97}.  An in-depth discussion of topologies in PSO is
beyond the scope of this paper.  The two most common topologies used in the
literature are the Ring topology and the Complete
topology~\cite{bratton-sis07}.  In the Ring topology, each particle has one
neighbor to either side of it; in the Complete topology, every particle is a
neighbor to every other particle.  In all topologies, a particle is also a
neighbor to itself, in that its own position is considered when updating the
particle's $\nbest$.  Thus with $p$ particles, using the Ring topology each
particle $i$ has three neighbors: $i-1$, $i$, and $i+1$.  With the Complete
topology, each particle has $p$ neighbors.  In this paper we use these
topologies, as well as a parallel adaptation of the Complete topology, called
Random, that has been shown to simulate the behavior of Complete with far less
communication~\cite{mcnabb-cec09}.  In the Random topology, each particle
randomly picks two other particles to share information with at each iteration,
along with itself.  Thus in both the Ring and the Random topologies, all
particles have 3 neighbors.
